Stack Safety as a Security Property

### 2020-07-05 main.tex
Description of the comservative policy

@@

\paragraph*{A conservative policy}

@@

>>>>

\newcommand*{\tagInstr}{\textsc{instr}}
\newcommand*{\tagCall}{\textsc{call}}
\newcommand*{\tagHa}{\textsc{h1}}
\newcommand*{\tagHb}{\textsc{h2}}
\newcommand*{\tagRa}{\textsc{r1}}
\newcommand*{\tagRb}{\textsc{r2}}
\newcommand*{\tagRc}{\textsc{r3}}
\newcommand*{\tagNoDepth}{\textsc{unused}}
\newcommand*{\tagStackDepth}[1]{\textsc{stack} ~ #1}
\newcommand*{\tagPCDepth}[1]{\textsc{pc} ~ #1}
\newcommand*{\tagSP}{\textsc{sp}}

@@

\paragraph*{A conservative policy}
%
The first micropolicy we present for eager stack safety is very similar to the
Depth Isolation policy presented by Roessler and DeHon \rb{cite}. Their policies
tag the stack memory with ownership information associated to each stack frame
(and to each separate object inside that frames, a more finely grained access
control that we do not need to consider), and registers holding stack pointers
with matching access permission information. Beyond those capabilities, the
compiler generates designated sequences of instructions with special tags that
grant them permission to manage the stack.

Our micropolicy works as follows. We assume for ease of exposition that no
arguments are passed on the stack \rb{Come back to this later?}. Initially, the
entire stack is tagged as unused with $\tagNoDepth$ \rb{actually, the ``initial
  frame'' should be tagged with its proper depth}, and the {\PCname} carries in
its tag at all times the depth of the current activation, starting with
$\tagPCDepth{0}$. The current activation depth needs to change exactly at those
points in the program where the contour of the stack safety property changes:

\begin{itemize}

\item From caller to callee, when the machine executes an instruction tagged as
  belonging to the call map with $\tagCall$. At this point, the {\PCname} depth
  is incremented to, say $\tagPCDepth{n}$; on function entry, the locations of
  the new frame is marked $\tagStackDepth{n}$.

\item From callee back to caller, when execution reaches the instruction
  immediately following the original call instruction and the $\SP$ is the same
  as at the time of that call. At this point, the {\PCname} is decremented;
  previously, the locations of the old frame are restored to $\tagNoDepth$.

\end{itemize}

To enforce this discipline, we define blessed header and return instruction
sequences. These are specific sequences of tagged instructions that are intended
to appear at the beginning and end of each function:

\[
  \begin{array}{c|c|c}
    {\sw ~ \rsp ~ \rra ~ 1} & {\tagHa, \tagInstr} & \text{Store RA in stack} \\
    {\addi ~ \rsp ~ \rsp ~ 2} & {\tagHb, \tagInstr} & \text{Increment SP by frame size}
  \end{array}
  \]

\[
  \begin{array}{c|c|c}
    {\lw ~ \rra ~ \rsp ~ -1} & {\tagRa, \tagInstr} & \text{Load RA} \\
    {\addi ~ \rsp ~ \rsp ~ -2} & {\tagRb, \tagInstr} & \text{Decrement SP by frame size} \\
    {\jalr ~ \rra ~ \rra} & {\tagRc, \tagInstr} & \text{Jump back}
  \end{array}
\]

\rb{Positioning of the above sequences, relation to running example. The tagging
  of stack frames would be considered part of these sequences.}

These sequences also enforce the $\SP$ discipline. When stack frames are of
fixed size, it suffices to increment $\SP$ and give it with a special tag in the
header sequence, which is later checked after decrementing $\SP$ during the exit
sequence to authorize the return. The well-formedness of the blessed sequences
can be checked statically. Only programs whose sequences are well-formed are
protected by the micropolicy, although programs are still protected (by
fail-stopping) even if the intended sequences are missing. By tracking the
execution of the blessed sequences in the {\PCname} tag state, the micropolicy
ensures that the entry sequence is executed correctly a declared call.

% If anything overwrites the SP in between, the exit sequence will fail, which is safe.

During normal execution, the {\PCname} uses the current activation depth in its
tag $\tagPCDepth{n}$ to gate load and store operations, which are allowed to
operate only on the corresponding stack frame, tagged $\tagStackDepth{n}$.
\rb{Note that we have also considered and implemented more flexible schemes
  based on allowing operations above the current activation depth. How do we put
  these together?}
\rb{Do we want to discuss variations involving frame pointers or more implicit
  return addresses?}
\rb{Where to talk about the role of the compiler?}

@@

\rb{The Static Authorities differs from the Depth Isolation policy in that frame
  ids aren't true activation ids, but unique ids associated to all activations
  of the same function. Discuss unique activation ids and compare with depth
  later, possibly when discussing lazy policies, as the difference becomes
  tangible there.}
